Recurrence of Geodesic Flow on the Necker Cube Surface

### 2017-08-28 paper.tex
082717

\newtheorem{thm}{Theorem}[section]
\newtheorem{Def}{Definition}[section]
\newtheorem*{thm*}{Theorem}
\newtheorem*{def*}{Definition}
\newtheorem{lem}{Lemma}[section]
\newtheorem*{rem}{Remark}

@@

\begin{Def} Identify $\mathbb C$ with $\mathbb{R}^2$ in the usual way. The closed path $\alpha\in\pi_1(\mathbf M, x_0)$ is the path contained in the maximal cylinder decomposition in direction $(1,1)$.
\end{Def}

@@

\begin{lem}
$\alpha$ is homotopic to the image of the closed geodesic in direction $(1,1)$ on $\mathbf{M}\backslash\text{Sing}(\mathbf M)$.
\begin{proof}
Let $\chi$ be a separatrix contained in either $C_1$ or $C_2$. Since a separatrix does not admit singularities, it is the image of a closed geodesic on $\mathbf{M}\backslash\text{Sing}(\mathbf M)$ with initial point $x_0$ on the strips of $C_1$ and $C_2$ with boundaries removed, denoted $C_1',$ $C_2'$. Express $[\alpha]$ as $\Sigma^{11}_{j=0}$ $\frac{1}{2}\gamma_j$ (a closed path climbing up the staircase). We show that the intersection numbers of $[\alpha]$ and $[\chi]$ are the same for every core cylinder curve $\gamma$, i.e. $i([\gamma_k], \Sigma^{11}_{j=0} \text{ } \frac{1}{2}[\gamma_j])=i([\gamma_k],[\chi])\text{ } \forall k=0,\dots,11$. And since $x_0$ is in either $C_1'$ or $C_2'$, $\alpha$ cannot climb up the staircase without intersecting the core curves. Thus, the sum of $\frac{1}{2}$ core curves of $[\alpha]$ have to have to intersect at least one other core curve.   \\
\textbf{Case one:} k is even. If k is even, then every curve $\gamma_k$ is oriented to the right. Since $\chi$ intersects every curve once, $i([\gamma_k],[\chi])=1$. No even indexed curves intersect eachother, so we need only consider when $j$ is odd. Now if $j$ is odd, it is incident (positively crossing) with only two horizontal curves, namely $\gamma_{j+1},\gamma_{j-1}$. Therefore $i([\gamma_k],[\alpha])=i([\gamma_{j-1}],\frac{1}{2}[\gamma_{j}])+i([\gamma_{j+1}],\frac{1}{2}[\gamma_{j}])=\frac{1}{2}(i([\gamma_{j-1}],[\gamma_{j}])+i([\gamma_{j+1}],[\gamma_{j}]))=\frac{1}{2}(1+1)=1.$\\
\textbf{Case two:} k is odd. If k is odd, then $[\chi]$ will have an intersection number of $-1$ with $[\gamma_k]$ since odd-indexed core curves are oriented upwards. Now since k is odd, we only consider when $j$ is even. Similarly, this means that $\gamma_j$ negatively intersects the two vertical core curves with adjacent indices. Hence, $i([\gamma_k],[\alpha])=i([\gamma_{j-1}],\frac{1}{2}[\gamma_{j}])+i([\gamma_{j+1}],\frac{1}{2}[\gamma_{j}])=\frac{1}{2}(i([\gamma_{j-1}],[\gamma_{j}])+i([\gamma_{j+1}],[\gamma_{j}]))=\frac{1}{2}(-1-1)=-1.$\\
We know intersection number to be a non-degenerate bilinear form, and if $\alpha$ and $\chi$'s abelianizations admit the same intersection numbers for every curve in the basis set of $H_1(\textbf M, \mathbb{Q})$, then $\alpha\sim\chi$.
\end{proof}
\end{lem}

@@

\begin{figure}[H]
\centering
\includesvg[width=1.8in]{cubesxyzpaths}
\includesvg[width=2.8in]{unfoldcutpath}
\end{figure}

@@

>>>>

\newtheorem{thm}{Theorem}[section]
\newtheorem{Def}{Definition}[section]
\newtheorem*{thm*}{Theorem}
\newtheorem*{def*}{Definition}
\newtheorem{lem}{Lemma}[section]
\newtheorem*{rem}{Remark}
\newtheorem*{conj}{Conjecture}

@@

\begin{Def} Identify $\mathbb C$ with $\mathbb{R}^2$ in the usual way. The closed path $\alpha\in\pi_1(\mathbf M, x_0)$ is the path contained in the maximal cylinder decomposition of $\mathbf M$ in direction $(1,1)$ with hol$(\alpha)=6+6i$.
\end{Def}

@@

\begin{lem}
$\alpha$ is homologous to the image of the closed geodesic in direction $(1,1)$ on $\mathbf{M}\backslash\text{Sing}(\mathbf M)$.
\begin{proof}
Let $\chi$ be a separatrix contained in either $C_1$ or $C_2$. Since a separatrix does not admit singularities, it is the image of a closed geodesic on $\mathbf{M}\backslash\text{Sing}(\mathbf M)$ with initial point $x_0$ on the strips of $C_1$ and $C_2$ with boundaries removed, denoted $C_1',$ $C_2'$. Express $[\alpha]$ as $\Sigma^{11}_{j=0}$ $\frac{1}{2}\gamma_j$ (a closed path climbing up the staircase). We show that the intersection numbers of $[\alpha]$ and $[\chi]$ are the same for every core cylinder curve $\gamma$, i.e. $i([\gamma_k], \Sigma^{11}_{j=0} \text{ } \frac{1}{2}[\gamma_j])=i([\gamma_k],[\chi])\text{ } \forall k=0,\dots,11$.    \\
\textbf{Case one:} k is even. If k is even, then every curve $\gamma_k$ is oriented to the right. Since $\chi$ intersects every curve once, $i([\gamma_k],[\chi])=1$. No even indexed curves intersect eachother, so we need only consider when $j$ is odd. Now if $j$ is odd, it is incident (positively crossing) with only two horizontal curves, namely $\gamma_{j+1},\gamma_{j-1}$. Therefore $i([\gamma_k],[\alpha])=i([\gamma_{j-1}],\frac{1}{2}[\gamma_{j}])+i([\gamma_{j+1}],\frac{1}{2}[\gamma_{j}])=\frac{1}{2}(i([\gamma_{j-1}],[\gamma_{j}])+i([\gamma_{j+1}],[\gamma_{j}]))=\frac{1}{2}(1+1)=1.$\\
\textbf{Case two:} k is odd. If k is odd, then $[\chi]$ will have an intersection number of $-1$ with $[\gamma_k]$ since odd-indexed core curves are oriented upwards. Now since k is odd, we only consider when $j$ is even. Similarly, this means that $\gamma_j$ negatively intersects the two vertical core curves with adjacent indices. Hence, $i([\gamma_k],[\alpha])=i([\gamma_{j-1}],\frac{1}{2}[\gamma_{j}])+i([\gamma_{j+1}],\frac{1}{2}[\gamma_{j}])=\frac{1}{2}(i([\gamma_{j-1}],[\gamma_{j}])+i([\gamma_{j+1}],[\gamma_{j}]))=\frac{1}{2}(-1-1)=-1.$\\
We know intersection number to be non-degenerate, and since $\alpha$ and $\chi$'s abelianizations admit the same intersection numbers for every curve in the spanning set of $H_1(\textbf M, \mathbb{Q})$,  $[\alpha]=[\chi]$.
\end{proof}
\end{lem}

@@

@@

\begin{conj}{\textbf{Periodicity of Geodesic Flow on the Necker cube surface.}}\\ Let $\theta$ be a direction obtained described as in Definition 3. Denote the geodesic flow on $\mathbf{S}$ by $F^t_{[\theta]}:T^1_{\bar{\Psi}(s)}(\mathbf{U}\backslash Sing(\mathbf U))\times\mathbb{R}_0^+\rightarrow\mathbf{S}$, where $F^t_{[\theta]}=\bar{\Psi}^{-1}(\bar{\Psi}(s)+t[\theta])$ for initial point $s\in\mathbf{S}$. Then the following is true:
\begin{enumerate}[label=(\roman*)]
\item If $\theta$ is rational and identified with an element in $\mathcal{O}$, then $F^t_{[\theta]}$ is periodic.
\item If $\theta$ is rational and identified with an element in $\mathcal{E}$, then $F^t_{[\theta]}$ is drift-periodic.
\end{enumerate}
\end{conj}

\newpage
\begin{lem}
The group $\mathbb{X}'$ acting on $\mathbb{Z}^2$ sends $\mathcal{O}$ to $\mathcal{O}$, and $\mathcal{E}$ to $\mathcal{E}$.
\begin{proof}
Since $\mathbb{X}'$ is generated by the elements $\mathbf{A}'$, $\mathbf{B}'$, and $\mathbf{R}'$, any matrix $G'\in\mathbb{X}'$ is of the form $G' = (\mathbf{A}')^{1_1}\circ(\mathbf{B}')^{1_2}\circ(\mathbf{R}')^{1_3}\circ(\mathbf{A}')^{2_1}\circ\dots(\mathbf{A}')^{n_1}\circ(\mathbf{B}')^{n_2}\circ(\mathbf{R}')^{n_3}$, where $i_k\in\mathbb{Z}$ for $i=1,\dots,n$ and $k=1,2,3.$ Let $x=\left(\begin{matrix}p \\ q  \end{matrix}\right),y\in\mathcal{O}$, and consider the equation $G'x=y$. Observe that $\left(\begin{matrix}1 && 2 \\ 0 && 1\end{matrix}\right)^l x=\left(\begin{matrix}p+2jq \\ q  \end{matrix}\right),$ and $ \left(\begin{matrix}1 && 0 \\ 2 && 1\end{matrix}\right)^m x=\left(\begin{matrix}p \\ q+2mp  \end{matrix}\right)$ for any $l,m\in\mathbb{Z}$. Also note that for any $j\in\mathbb{Z}$, $(\mathbf{R}')^{m}x=\left(\begin{matrix}p \\ q  \end{matrix}\right),\left(\begin{matrix}-q\\ p  \end{matrix}\right),\left(\begin{matrix}-p \\ -q  \end{matrix}\right),\left(\begin{matrix}q \\ -p  \end{matrix}\right)$ when $j \mod{4}\equiv0,1,2,3$, respectively. In any case, the product of any power of a generator of $\mathbb{X}'$ and any $x\in\mathcal{O}$ is an element of $\mathcal{O}$ . By letting $l=i_1,m=i_2,$ and $j=i_3$, we first consider the base case when $i=n$. Let $G'=G'_1\circ\dots\circ G'_n$, such that $G'_i=(\mathbf{A}')^{i_1}\circ(\mathbf{B}')^{i_2}\circ(\mathbf{R}')^{i_3}$. Since $n_1,n_2,n_3$ are arbitrary integers, $G'_n x \in\mathcal{O}$. Suppose for some $b< n-1$, $G'_{n-b}\circ\dots\circ G'_n x =y'\in\mathcal{O}$. Therefore $y'=(G'_1\circ\dots\circ G'_{b})^{-1}y$, which implies that $(G'_1\circ\dots\circ G'_{b})^{-1}$ preserves the set $\mathcal{O}$. Otherwise, if $y\in\mathcal{E}$, there exists at least one $G'_i$ for $1< i< b$ and $\tau\in\mathcal{E}$ such that $G'^{-1}_i \tau = (\mathbf{R}')^{-i_3}\circ(\mathbf{B}')^{-i_2}\circ(\mathbf{A}')^{-i_1} \tau \in \mathcal{O}$, a contradiction. Since elements in $\mathbb{X}'$ are invertible, $G'_1\circ\dots\circ G'_{b}$ must also preserve $\mathcal{O}$. By inductive reasoning, we conclude that  $G'_1\circ\dots\circ G'_n x = G' x = y$. $x\in\mathcal{O}$ if and only if $y\in\mathcal{O}$ since $G'$ is invertible.\\
Proving that $G'(\mathcal{E})=\mathcal{E}$ is done in the same way.
\end{proof}
\end{lem}
